A Parallel General Purpose Multi-Objective Optimization Framework, with Application To Beam Dynamics

### 2018-03-15 paper/paper.tex
added sentence on optimizer being in OPAL
starting to add awa parameters

With access to the \textsc{OPAL} forward solver the framework is able to
  tackle a multitude of optimization problems arising in the domain of
  particle accelerators.
By abiding to the API from Listing 3 it becomes trivial to add new forward
  solvers to the framework (see Appendix A).
If the objectives and constraints are simple arithmetical expressions, the
  \texttt{FunctionEvaluator} simulator can be used.
Using functors and the default expression primitives already powerful
  multi-objective optimization problems can be specified, i.e.\ the benchmark
  problem presented in \cite{hbwh:05}:
%
\begin{align} \label{eqn:bench}
  \text{ min} \quad & \quad \left[ 1 - \exp \left( -1 \left(
    \left(x_1 - \frac{1}{\sqrt{3}} \right)^2 +
    \left(x_2 - \frac{1}{\sqrt{3}} \right)^2 +
    \left(x_3 - \frac{1}{\sqrt{3}} \right)^2 \right)\right), \right. \\
                    & \quad \left. 1 - \exp \left( -1 \left(
    \left(x_1 + \frac{1}{\sqrt{3}} \right)^2 +
    \left(x_2 + \frac{1}{\sqrt{3}} \right)^2 +
    \left(x_3 + \frac{1}{\sqrt{3}} \right)^2 \right)\right) \right]^T \nonumber \\
    \text{s.t.} \quad & \quad -1 \le x_i \le 1, \quad i=1,2,3 \nonumber
  \text{.}
\end{align}
%
Using the default expression primitives this can be stated in an input file
  as:

@@

The authors thank the \textsc{SwissFEL} and AWA teams for contributing to the
  formulation of optimization problems. We gratefully acknowledge the computing resources provided on Bebop, a high-performance computing cluster operated by the Laboratory Computing Resource Center at Argonne National Laboratory.
\appendix

>>>>

With access to the \textsc{OPAL} forward solver the framework is able to
  tackle a multitude of optimization problems arising in the domain of
  particle accelerators.
  The framework is also integrated into \textsc{OPAL} so that users can 
  define optimization problems within an input file, requiring no 
  additional knowledge or installation of the API to use it.

By abiding to the API from Listing 3 it becomes trivial to add new forward
  solvers to the framework (see Appendix A).
If the objectives and constraints are simple arithmetical expressions, the
  \texttt{FunctionEvaluator} simulator can be used.
Using functors and the default expression primitives already powerful
  multi-objective optimization problems can be specified, i.e.\ the benchmark
  problem presented in \cite{hbwh:05}:
%
\begin{align} \label{eqn:bench}
  \text{ min} \quad & \quad \left[ 1 - \exp \left( -1 \left(
    \left(x_1 - \frac{1}{\sqrt{3}} \right)^2 +
    \left(x_2 - \frac{1}{\sqrt{3}} \right)^2 +
    \left(x_3 - \frac{1}{\sqrt{3}} \right)^2 \right)\right), \right. \\
                    & \quad \left. 1 - \exp \left( -1 \left(
    \left(x_1 + \frac{1}{\sqrt{3}} \right)^2 +
    \left(x_2 + \frac{1}{\sqrt{3}} \right)^2 +
    \left(x_3 + \frac{1}{\sqrt{3}} \right)^2 \right)\right) \right]^T \nonumber \\
    \text{s.t.} \quad & \quad -1 \le x_i \le 1, \quad i=1,2,3 \nonumber
  \text{.}
\end{align}
%
Using the default expression primitives this can be stated in an input file
  as:

@@

The authors thank the \textsc{SwissFEL} and AWA teams for contributing to the
  formulation of optimization problems. We gratefully acknowledge the computing resources provided on Bebop, a HPC cluster operated by the LCRC at ANL.
\appendix

### 2018-03-15 paper/theor-experiment.tex
starting to add awa parameters

\subsection{Ferrario Matching Point}

@@

\subsection{AWA}
Next we apply the optimization framework to an ongoing project in accelerator physics.
The model is taken from the Argonne Wakefield Accelerator facility, which 
houses two rf photoinjector beam lines. We consider the high charge beam line
that consists of an rf photocathode gun, two solenoids, six linear accelerating cavities, 
and

>>>>

\subsection{Ferrario Matching Point}\label{ferrario}

@@

\subsection{AWA Photoinjector Optimization}
Next we apply the optimization framework to the high charge beam line at the AWA, 
which consists of an rf photocathode gun, two solenoids, six linear accelerating cavities, 
followed by four quadrupoles. This beam line has served as the injector for Two Beam 
Acceleration (TBA) experiments in the past [XX], and will be used in the future for 
similar experiments. One of the main constraints on prior experimental 
results was the beam size and emittance as the beam passed through small aperture 
wakefield structures. In an attempt to maximize charge transmission in future experiments
we optimize several parameters in the gun and quadrupole leading into the 
TBA section of the beam line.

Compared to the Ferrario experiment in Section \ref{ferrario}, 
this problem is larger by a factor of two in both design variables and objectives.
The problem is also less predictable, meaning there is not a clear idea
before hand where the pareto front will fall. This work is also meaningful in that
it will guide future operations at the AWA.

The design variables include both solenoid current, IBF and IM,
the gun phase, GPHASE, and the laser full width half maximum in 
the longitudinal direction. The definition of the design variables 
was included in the OPAL input file:
\vspace{0.2cm}
{\footnotesize \begin{verbatim}
// Gun variables
dv0: DVAR, VARIABLE="IBF",    LOWERBOUND=350.0, UPPERBOUND=500.0;
dv1: DVAR, VARIABLE="IM",     LOWERBOUND=170.0, UPPERBOUND=260.0;
dv2: DVAR, VARIABLE="GPHASE", LOWERBOUND=-30.0, UPPERBOUND=0.0;
dv3: DVAR, VARIABLE="FWHM",   LOWERBOUND=1.5,   UPPERBOUND=10.0;

// Quad variables
dv4: DVAR, VARIABLE="KQ1", LOWERBOUND=-8.0, UPPERBOUND=8.0;
dv5: DVAR, VARIABLE="KQ2", LOWERBOUND=-8.0, UPPERBOUND=8.0;
dv6: DVAR, VARIABLE="KQ3", LOWERBOUND=-8.0, UPPERBOUND=8.0;
dv7: DVAR, VARIABLE="KQ4", LOWERBOUND=-8.0, UPPERBOUND=8.0;

	\end{verbatim}}
\vspace{0.2cm}

 \vspace{0.2cm}
 {\footnotesize \begin{verbatim}
rmss1:  OBJECTIVE,EXPR="fabs(statVariableAt('rms_s',16.45))";
rmsx1:  OBJECTIVE,EXPR="fabs(statVariableAt('rms_x',16.45))";
emitx1: OBJECTIVE,EXPR="fabs(statVariableAt('emit_x',16.45))";

rmss2:  OBJECTIVE,EXPR="fabs(statVariableAt('rms_s',18.45))";
rmsx2:  OBJECTIVE,EXPR="fabs(statVariableAt('rms_x',18.45))";
emitx2: OBJECTIVE,EXPR="fabs(statVariableAt('emit_x',18.45))";

de1:    OBJECTIVE,EXPR="fabs(statVariableAt('dE',16.45))";

	\end{verbatim}}
\vspace{0.2cm}

All simulations for this experiment were carried out on Bebop a
high performance computing (HPC)
cluster provided by the Laboratory Computing Resource Center (LCRC)
at Argonne National Laboratory (ANL). We used Intel Knights Landing 
(KNL) processors at 1.3 GHz with 128 GB of memory per node, 
and 64 cores per node. There are 352 compute nodes available on 
Bebop, with a total of 22528 cores. 
This allows for very large optimization jobs, like the AWA case.

OPAL-T was used as the forward solver, and a study on time step and 
number of particles was done to reduce the time to simulation while 
maintaining the physics of interest. We choose ten thousand particles 
in combination with adjusting the time step near sensitive elements (i.e. quadurpoles). 
The resulting simulations are low fidelity, but closely approximate 
the mid fidelity simulations for metrics of interest, as shown in Fig. \ref{tstep}.
\begin{figure}\ref{tstep}
	\centering
	%\includegraphics[width=0.8\linewidth]{Report/}
	\caption{Comparison of different fidelity models.}
	\label{tstep}
\end{figure}

As in Section \ref{ferrario}, 
we also have to define several beam line parameters that 
are not included in the optimization design variables above.
These
